---
title: "Coursework 1 2020"
author: "George Hilton"
date: "30/11/2020"
output:
  html_notebook: default
  html_document:
    df_print: paged
  df_print: paged
  pdf_document: default  
---


Using the optimize function to efficiently calculate M as well as the paramaters a and b for the beta distribution.

```{r}

```

```{r include=FALSE}
library(ggplot2, microbenchmark)
```

Optimisation to find M, a, b. Construct f*, and f*/g

```{r include=FALSE}
foverg <- function(x, a, b){
  ((1/(x*(1-x)))*exp((-1/9)*(0.3+log(x/(1-x)))^2))/((x^(a-1)*(1-x)^(b-1)*factorial(a+b-1))/(factorial(a-1)*factorial(b-1)))
}

optimiseM <- function(params){
  optimize(foverg, a = params[1], b=params[2], lower=0, upper=1, maximum=TRUE)$objective}

results <- optim(par = c(0.7, 0.8),optimiseM, lower = 0, upper = Inf, method = "L-BFGS-B")

a <- results$par[1]
b <- results$par[2]
M <- results$value

fovergP <- function(x){
  ((1/(x*(1-x)))*exp((-1/9)*(0.3+log(x/(1-x)))^2))/((x^(a-1)*(1-x)^(b-1)*factorial(a+b-1))/(factorial(a-1)*factorial(b-1)))
}


fstar <- function(x){
  (1/(x*(1-x)))*exp((-1/9)*(0.3+log(x/(1-x)))^2)
}

g <- function(x){
  (x^(a-1)*(1-x)^(b-1)*factorial(a+b-1))/(factorial(a-1)*factorial(b-1))
}
```

plot1, plotting of f* and mg

```{r}
knitr::opts_chunk$set(echo=FALSE)
plot1 <- function(){
  x=seq(0.1,0.99,l=100)
df = data.frame(x=x, pdf1=fstar(x), pdf2=M*g(x))
mylabels=list(expression(f[X](x)),expression(Mg[Y](x)))

#plotting the graph of the pdf and its envelope function
p <- ggplot(df)
p + geom_line(aes(x,pdf1,colour="pdf1"))+
  geom_line(aes(x,pdf2,colour="pdf2"))+ 
  labs(y="Density", title=expression("Comparison of "~f[X](x)~"and "~Mg[X](x)))+
  scale_colour_manual("", values=c("pdf1"="#CC3333","pdf2"="#3399CC"), labels=mylabels)
}

plot1()
```

construction of wl and wu

```{r}
knitr::opts_chunk$set(echo=FALSE)
sbox <- function(){
  # finding a value for the top of my rectangle. i will use the minimum of the dip 
top = optimize(fovergP, lower=0.25, upper=0.75, maximum=FALSE)$objective
# finding the left and right values for my rectangle
transformDown <- function(x, transform){fovergP(x)-transform}
# 0.01 is a reasonable lower bound as i checked fovergP(0.01) < top
leftVal = uniroot(transformDown, transform=top, lower=0.01, upper=0.5)$root
# 0.99 also reasonable upper bound as fovergP(0.99) < top
rightVal = uniroot(transformDown, transform=top, lower=0.5, upper=0.99)$root
return(c(top, leftVal, rightVal))
}

bbox <- function(){
  # finding a value for the top of my rectangle. i will use the minimum of the dip 
top = optimize(fovergP, lower=0, upper=1, maximum=TRUE)$objective
return(top)
}

svals = sbox()
bval = bbox()

wl <- function(y){ifelse(svals[2] < y & y < svals[3], svals[1], 0)}
wu <- function(y){ifelse(0 < y & y < 1, bval)}

```

plot of wl, wu and f/g

```{r}
knitr::opts_chunk$set(echo=FALSE)
plot2 <- function(){
  x=seq(0.0000001,0.9999999,l=100)
  df2 = data.frame(x=x, pdf1=fovergP(x), pdf2=wl(x), pdf3=wu(x))
  mylabels=list(expression(f/g[X](x)), expression(W_L(x)), expression(Wu(x)))

  p <- ggplot(df2)
  p + geom_line(aes(x,pdf1,colour="pdf1"))+
  geom_line(aes(x,pdf2,colour='pdf2'))+
  geom_line(aes(x,pdf3,colour='pdf3'))+
  labs(y="Density", title=expression("Plot of "~f/g[X](x)~"and"~W_L(x)))+
  scale_colour_manual("", values=c("pdf1"="#CC3333", "pdf2"="#3399CC", 'pdf3'='#336600'), labels=mylabels)
}

plot2()
```
integrals to compute normalising constant, and theoretical acceptance probability
```{r}
integrand1 <- function(x){ #theoretical acceptance probability
 ((1/(x*(1-x)))*exp((-1/9)*(0.3+log(x/(1-x)))^2))/M
}

integrand2 <- function(x){ #theoretical acceptance probability
 ((1/(x*(1-x)))*exp((-1/9)*(0.3+log(x/(1-x)))^2))
}

theta = integrate(integrand1, lower=0, upper=1)
theta = theta$value
N = integrate(integrand2, lower=0, upper=1)
N = N$value
f <- function(x){
  1/N * (1/(x*(1-x)))*exp((-1/9)*(0.3+log(x/(1-x)))^2)
}

```

generating samples using squeeze and without

```{r}

rdsSQUEEZE <- function(n){ #the function which generates the sample
  x <- vector("numeric") #vector to keep track of simulated values 
  count <- 0
  total <- 0
  while(count<n){
    ngen = ceiling(1/theta * (n-count))
    u <- runif(ngen) #generate u, uniformly, for accept/reject
    v <- runif(ngen) #generate v, uniformly, for use in the envelope function
    y <- qbeta(v, a, b) #generate Exp(1) using inversion
    out <- ifelse(M*u <= wl(y), T, ifelse(M*u > wu(y), F, ifelse(u > fovergP(y)/M, F, T))) 
    x <- c(x, y[out]) #accept X=y if u <= f(y)/Mg(y)
    count <- length(x) #the next two lines are used to keep track of the test and acceptance
    total <- total + ngen
 }
 predictedp = format(theta) #theoretical acceptance probability
 actualp = format(count/total) #actual acceptance probability
 return(list(x=x, 'Predicted Acceptance Probability'=predictedp, 'Actual Acceptance Probability'=actualp))
}


rds <- function(n){ #the function which generates the sample
  x <- vector("numeric") #vector to keep track of simulated values 
  count <- 0
  total <- 0
  while(count<n){
    ngen = ceiling(1/theta * (n-count))
    u <- runif(ngen) #generate u, uniformly, for accept/reject
    v <- runif(ngen) #generate v, uniformly, for use in the envelope function
    y <- qbeta(v, a, b) #generate Exp(1) using inversion
    x <- c(x, y[u <= fovergP(y)/M]) #accept X=y if u <= f(y)/Mg(y)
    count <- length(x) #the next two lines are used to keep track of the test and acceptance
    total <- total + ngen
 }
 predictedp = format(theta) #theoretical acceptance probability
 actualp = format(count/total) #actual acceptance probability
 return(list(x=x, 'Predicted Acceptance Probability'=predictedp, 'Actual Acceptance Probability'=actualp))
}

```

benchmarking

```{r}

benchmark <- function(n, t){
  results <- microbenchmark(rdsSQUEEZE(n), rds(n), times = t)
  return(results)
}

benchmark(100000,1)


```


plot of f against data (need to add for both rds funcs)

```{r}
plot3 <-function(){
  n = 1000 #sample size
  x = rds(n)$x #ensuring we'll have numeric data
  df1 = data.frame(x, pdf=f(x))
  df2 = data.frame(x)
  mylabels=list(expression(f[X](x)))
  #plotting the graph
  p <- ggplot(df1)
  p + geom_histogram(aes(x,y= ..density..), breaks=seq(0, 1,l=100))+
  geom_line(aes(x,pdf,colour="pdf"))+ 
  labs(y="density", title=expression("Comparison of Histogram and "~f[X](x)))+
  scale_colour_manual("", values=c("pdf"="red"), labels=mylabels)
}

plot3()

```

diagnostic plots

```{r}









```